---
title: "Class_12_RNA_seq_021624"
author: "Johann Tailor"
format: pdf
editor: visual
---


Reading the sample summary file into R:

```{r}
expression_file <- read.table("https://bioboot.github.io/bggn213_W24/class-material/rs8067378_ENSG00000172057.6.txt")


#Finding the summary of the file loaded:
summary(expression_file)

```

Identifying how many patients are there in each genotype category:
```{r}
All_AG <- grep("A/G", expression_file$geno)
All_GG <- grep("G/G", expression_file$geno)
All_AA <- grep("A/A", expression_file$geno)

All_AG





#A/G_patients <- row.names(df[All_AG, ])
```
Another way:

Trying codes to figure it out how to get row.names from the grepED files.
```{r}
matching_AG <- rownames(expression_file)[grep("A/G", expression_file$geno)]
matching_GG <- rownames(expression_file)[grep("G/G", expression_file$geno)]
matching_AA <- rownames(expression_file)[grep("A/A", expression_file$geno)]

matching_GG
matching_AA
matching_AG
```

> Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

Here are the median for each genotype:
```{r}
summary(expression_file[expression_file[,2] == "A/A",3] )
summary(expression_file[expression_file[,2] == "G/G",3] )
summary(expression_file[expression_file[,2] == "A/G",3] )


```
ANSWER:

Median for each genotype is as follows


A/A : 31.25
G/G : 20.074
A/G : 25.065

>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

Let's generate a box plot to compare the means of the three genotypes and their expression of ORMDL3 gene:

```{r}


box_plot_genotypes <- boxplot(exp~geno, data=expression_file, xlab="rs8067378 genotype", ylab="ENSG00000172057.4 (RPKM)")

box_plot_genotypes
```

ANSWER:

I do think the SNP G|G renders overall less gene expression of ORMDL3 compared to A|A; the mean of A|A is 31.82 and G|G is 20.5, respectively. 
